Mixed language processing increases cross-language phonetic transfer in Bengali-English Bilinguals

### 2021-06-08 draft_glossa-psych/glossa-psych_AM-ID_manuscript.tex
abstract

% \author[Auromita \& Indranil]% short form of the author names for the running header. If no short author is given, no authors print in the headers.
% {%as many authors as you like, each separated by \AND.
%   \spauthor{Auromita Mitra\\ 
%   \institute{The EFL University}\\
%   \small{%105, Bd. Raspail, 75005 Paris\\
%   auromita.mitra@gmail.com}
%   }
%   \AND
%   \spauthor{Indranil Dutta \\
%   \institute{Jadavpur University}\\
%   \small{%Warmoesberg 26, 1000 Brussel\\
%   indranildutta.lnl@jadavpuruniversity.in}
%   }%
% }
%\usepackage{gb4e}
\begin{document}

@@

Word count: 6944

\begin{abstract}
Existing research demonstrates that mixed-language productions of bilingual speakers differ from productions in a single language. Studies have largely focused on temporal properties of consonants, and reported varied, and often contradictory, results. The present study examines the effect of mixed-language processing on two L2 vowel categories in a group of highly proficient Bengali-English bilingual speakers in India. Following from a discussion in \citep{olson2013bilingual}, we compare productions elicited in two paradigms-- cued picture-naming and code-switching. Results indicate a dynamic increase in cross-language phonetic transfer during mixed-language use leading to a shift towards L1 norms, asymmetrical shifts in the two categories, and no difference in transfer patterns between the two paradigms. 
\end{abstract}

@@

Bi/multilingual speakers can distinguish between the phonetic norms of their languages \cite{caramazza1973acquisition,macleod2010impact,bosch2003simultaneous} and therefore maintain separate sound categories for each language. However, these categories are not autonomous-- they show cross-language influence in both perception and production \cite{flege1995second,fowler2008cross,flege2002assessing}. What causes, affects, and results from such influence? Answers to these questions provide important insights into how language `systems' are represented and processed in the mind. %Given that a majority of the world's population is multilingual, this is a central concern for any realistic theory of language.
%show cross-language-- rephrase

%keep just the types here, and put the next 2 lines in the causes&effects section?
Cross-language influence at the level of sounds can be studied broadly in two kinds of conditions:
(i) While a bilingual speaker is operating in any one of their languages, by comparing bilingual speech to monolingual norms \citep[e.g.][]{guion2003vowel,caramazza1973acquisition,flege1987production}; %These studies often compare the speech of bilinguals to monolingual norms, and view the influence as changes to long-term memory representations as a result of acquiring an L2 \citep{guion2003vowel,caramazza1973acquisition,flege1987production}; 
(ii) When both languages of a bilingual speaker are co-activated, by comparing bilingual speakers' mixed-language speech to their own norms while using a single language \citep[e.g.][]{grosjean1994going, bullock2009trying,elias2017effects, simonet2014phonetic}. %These compare productions during mixed-language use to participants' own productions while using a single language. The resulting influence is variously thought to involve short-term memory, online processing costs, language mode, and context-awareness. 
The present study is concerned with the latter. In production, researchers have variously referred to this kind of influence as transfer, drift, accommodation, and interference-- these terms are used interchangeably here to indicate any interaction between two sets of phonetic norms.

@@

Considering these facts, the present study focuses on short-term phonetic interaction during mixed-language use for the following reasons:

@@

\subsection{Duration: How long do effects last?} \label{duration}

>>>>

% \author[Auromita \& Indranil]% short form of the author names for the running header. If no short author is given, no authors print in the headers.
% {%as many authors as you like, each separated by \AND.
%   \spauthor{Auromita Mitra\\ 
%   \institute{The EFL University}\\
%   \small{%105, Bd. Raspail, 75005 Paris\\
%   auromita.mitra@gmail.com}
%   }
%   \AND
%   \spauthor{Indranil Dutta \\
%   \institute{Jadavpur University}\\
%   \small{%Warmoesberg 26, 1000 Brussel\\
%   indranildutta.lnl@jadavpuruniversity.in}
%   }%
% }
%\usepackage{gb4e}

\begin{document}

@@

Word count: 7627

\begin{abstract}
This study investigates the phonetic outcome of mixed-language production in a group of highly proficient bilingual speakers of Bengali and English, to better understand the nature of cross-language phonetic interaction. Existing research has largely focused on a small set of phonologically related languages and used voice onset time (VOT) to measure transfer. Results show that transfer is mediated by both language-specific features and linguistic experience of participants, pointing to an urgent need for diversifying the range of languages and populations studied. We compare productions from a yet-unstudied population in two conditions—unilingual L2 (English) and mixed (English and Bengali). The spectral properties of two L2 vowel categories— \nt{\ae}(common to L1 and L2) and \nt{2} (absent in L1) are examined as sites for transfer. To understand how this unfolds dynamically, we measure F1 and F2 at five points over the course of articulating the vowel. We find a dynamic change in both vowels during mixed production, with \nt{2} shifting towards a related L1 category. We interpret this as a temporary increase in cross-language phonetic influence during mixed-language use, leading to a shift towards L1 norms. asymmetrical shifts in the two categories, and no difference in transfer patterns between the two paradigms. 
\end{abstract}

@@

Bi/multilingual speakers can distinguish between the phonetic norms of their languages and maintain separate sound categories for each language \citep{caramazza1973acquisition,macleod2010impact,bosch2003simultaneous}. However, these categories are not autonomous-- they influence each other across languages in both perception and production \citep[e.g.][]{flege1995second,fowler2008cross,flege2002assessing}, and the nature of such interaction provides crucial insights into how language `systems' are cognitively represented and processed. This study reports on the phonetic outcomes of mixed-language production in proficient bilingual speakers of Bengali and English.\\%Given that a majority of the world's population is multilingual, this is a central concern for any realistic theory of language.
%show cross-language-- rephrase

%keep just the types here, and put the next 2 lines in the causes&effects section?
Cross-language influence at the level of sounds can be studied broadly in two kinds of conditions:
(i) While a bilingual speaker is operating in any one of their languages, by comparing bilingual speech to monolingual norms \citep[e.g.][]{guion2003vowel,caramazza1973acquisition,flege1987production}; %These studies often compare the speech of bilinguals to monolingual norms, and view the influence as changes to long-term memory representations as a result of acquiring an L2 \citep{guion2003vowel,caramazza1973acquisition,flege1987production}; 
(ii) When both languages of a bilingual speaker are co-activated, by comparing speakers' mixed-language speech to their own norms while using a single language \citep[e.g.][]{grosjean1994going, bullock2009trying,elias2017effects, simonet2014phonetic}. %These compare productions during mixed-language use to participants' own productions while using a single language. The resulting influence is variously thought to involve short-term memory, online processing costs, language mode, and context-awareness. 
The present study is concerned with the latter. In production, researchers have variously referred to this kind of influence as transfer, drift, accommodation, and interference-- these terms are used interchangeably here to indicate any interaction between two sets of phonetic norms.\\

@@

@@

\subsection{Duration of effects} \label{duration}
